Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists

### 2024-11-13 paper.Rmd
correcting image captions + reference to repository

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
# read data
pre_surveys <- readRDS("data/pre_workshop_surveys.rds")
post_surveys <- readRDS("data/post_workshop_surveys.rds")
# factore levels for charts
factor_levels <- c("Excellent", "Good", "Sufficient", "Insufficient", "Not shared")
# readable event names
event_names <- as_labeller(c("202304_CAA" = "CAA", "202308_EAA" = "EAA", "202310_CAA-DE-NLFl" = "CAA-DE-NLFl", "202311_Reuvensdagen" = "Reuvensdagen", "202311_CAA_UK" = "CAA_UK", "202312_Leiden" = "Leiden", "202401_Aarhus" = "Aarhus"))
events <- read.csv("data/events.csv", sep = ";")

@@

The surveys were analysed using R [@rcoreteam2023]. The following packages were used for analyses: ggplot2 [@wickham2016], dplyr [@wickham2023a], tidyr [@wickham2023], forcats[@wickham2023b], lubridate [@grolemund2011] and stringr [@wickham2023c]. For the sake of reproducibility [@marwick2017] this paper was written in R Markdown (<https://rmarkdown.rstudio.com/>), with data and code available at <https://github.com/ABMArchaeologists/ABMA_paper> (... insert Zenodo reference...)

@@

```{r, abm-knowledge, fig.cap="Respondents knowledge of ABM per event knowledge of the various softwares for ABM, before participating in the workshops.", fig.show="hold", out.width="50%", echo=FALSE}
pre_surveys %>%
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>%
  select(Do.you.know.what.Agent.Based.Modeling.is., Workshop) %>%
  group_by(`Do.you.know.what.Agent.Based.Modeling.is.`, Workshop) %>%
  mutate(count_response = n()) %>%
  distinct() %>%
  ggplot(aes(fill = `Do.you.know.what.Agent.Based.Modeling.is.`, x=Workshop, y = count_response)) +
  geom_bar(na.rm = TRUE, position="fill", stat="identity") +
  geom_text(aes(label=count_response), position = position_fill(vjust = 0.5), size = 2.5, colour = "grey41") + scale_x_discrete(labels = c("CAA", "EAA", "CAA-DE-NLFl", "Reuvensdagen", "CAA_UK", "Leiden", "Aarhus")) +
  xlab("") + scale_fill_discrete(name = "ABM known") + ylab("Percentage") +  scale_y_continuous(labels = scales::label_percent(scale = 100, prefix = "", suffix = "")) + coord_flip()
pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
    mutate(count_software = n()) %>%
  ggplot(aes(x=reorder(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`, count_software))) + geom_bar() + scale_x_discrete(labels = ) + xlab("ABM Software known") + coord_flip()
```

@@

```{r, rating-teaching, fig.cap="Respondents rating of the teaching material in general faceted for each event.", fig.width=11, fig.show='hold', out.height="50%", echo=FALSE}
post_surveys %>%
  mutate(How.would.you.rate.the.teaching.material. = fct_relevel(How.would.you.rate.the.teaching.material., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.teaching.material.)) + geom_bar() +
  xlab("Rating of the teaching material") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
post_surveys %>%
  mutate(How.would.you.rate.the.trainers.teachers. = fct_relevel(How.would.you.rate.the.trainers.teachers., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.trainers.teachers.)) + geom_bar() +
  xlab("Rating of the teacher") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
library(ggpubr)
# read data
pre_surveys <- readRDS("data/pre_workshop_surveys.rds")
post_surveys <- readRDS("data/post_workshop_surveys.rds")
# factore levels for charts
factor_levels <- c("Excellent", "Good", "Sufficient", "Insufficient", "Not shared")
# readable event names
event_names <- as_labeller(c("202304_CAA" = "CAA", "202308_EAA" = "EAA", "202310_CAA-DE-NLFl" = "CAA-DE-NLFl", "202311_Reuvensdagen" = "Reuvensdagen", "202311_CAA_UK" = "CAA_UK", "202312_Leiden" = "Leiden", "202401_Aarhus" = "Aarhus"))
events <- read.csv("data/events.csv", sep = ";")

@@

The surveys were analysed using R [@rcoreteam2023]. The following packages were used for analyses: ggplot2 [@wickham2016], dplyr [@wickham2023a], tidyr [@wickham2023], forcats[@wickham2023b], lubridate [@grolemund2011] and stringr [@wickham2023c]. For the sake of reproducibility [@marwick2017] this paper was written in R Markdown (<https://rmarkdown.rstudio.com/>), with data and code available at <https://github.com/ABMArchaeologists/ABMA_paper> [@visser].

@@

```{r, abm-knowledge, fig.cap="Respondents knowledge of ABM per event (top), and knowledge of the various softwares for ABM (bottom), before participating in the workshops.", echo=FALSE}
knowledge_plot <- pre_surveys %>%
  filter(`Do.you.know.what.Agent.Based.Modeling.is.`!="Not shared") %>%
  select(Do.you.know.what.Agent.Based.Modeling.is., Workshop) %>%
  group_by(`Do.you.know.what.Agent.Based.Modeling.is.`, Workshop) %>%
  mutate(count_response = n()) %>%
  distinct() %>%
  ggplot(aes(fill = `Do.you.know.what.Agent.Based.Modeling.is.`, x=Workshop, y = count_response)) +
  geom_bar(na.rm = TRUE, position="fill", stat="identity") +
  geom_text(aes(label=count_response), position = position_fill(vjust = 0.5), size = 2.5, colour = "grey41") + scale_x_discrete(labels = c("CAA", "EAA", "CAA-DE-NLFl", "Reuvensdagen", "CAA_UK", "Leiden", "Aarhus")) +
  xlab("") + scale_fill_discrete(name = "ABM known") + ylab("Percentage") +  scale_y_continuous(labels = scales::label_percent(scale = 100, prefix = "", suffix = "")) + coord_flip()
software_plot <- pre_surveys %>%
  select(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
  filter(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`!="") %>%
  separate_rows(Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice, sep = ',\\s*') %>%
  group_by(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`) %>%
    mutate(count_software = n()) %>%
  ggplot(aes(x=reorder(`Which.software.do.you.know.for.Agent.Based.Modeling....Selected.Choice`, count_software))) + geom_bar() + scale_x_discrete(labels = ) + xlab("ABM Software known") + coord_flip()
ggarrange(knowledge_plot, software_plot, nrow = 2, ncol = 1)
```

@@

```{r, rating-teaching, fig.cap="Respondents rating of the teaching material (top) and teachers (bottom) in general faceted for each event.", fig.width=11, fig.height=8, echo=FALSE}
rate_material_plot <- post_surveys %>%
  mutate(How.would.you.rate.the.teaching.material. = fct_relevel(How.would.you.rate.the.teaching.material., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.teaching.material.)) + geom_bar() +
  xlab("Rating of the teaching material") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
rate_teacher_plot <- post_surveys %>%
  mutate(How.would.you.rate.the.trainers.teachers. = fct_relevel(How.would.you.rate.the.trainers.teachers., factor_levels)) %>%
  ggplot(aes(x=How.would.you.rate.the.trainers.teachers.)) + geom_bar() +
  xlab("Rating of the teacher") + facet_grid(~Workshop, labeller = event_names) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
ggarrange(rate_material_plot, rate_teacher_plot, nrow = 2, ncol = 1)
```

### 2024-11-13 paper.tex
correcting image captions + reference to repository

\title{Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists}
\author{}
\date{\vspace{-2.5em}November 12, 2024}

@@

The surveys were analysed using R (R Core Team 2023). The following packages were used for analyses: ggplot2 (Wickham 2016), dplyr (Wickham et al. 2023), tidyr (Wickham, Vaughan and Girlich 2023), forcats(Wickham 2023a), lubridate (Grolemund and Wickham 2011) and stringr (Wickham 2023b). For the sake of reproducibility (Marwick 2017) this paper was written in R Markdown (\url{https://rmarkdown.rstudio.com/}), with data and code available at \url{https://github.com/ABMArchaeologists/ABMA_paper} (\ldots{} insert Zenodo reference\ldots)

@@

\begin{figure}
\includegraphics[width=0.5\linewidth]{paper_files/figure-latex/abm-knowledge-1} \includegraphics[width=0.5\linewidth]{paper_files/figure-latex/abm-knowledge-2} \caption{Respondents knowledge of ABM per event knowledge of the various softwares for ABM, before participating in the workshops.}\label{fig:abm-knowledge}
\end{figure}

@@

\begin{figure}
\includegraphics[height=0.5\textheight]{paper_files/figure-latex/rating-teaching-1} \includegraphics[height=0.5\textheight]{paper_files/figure-latex/rating-teaching-2} \caption{Respondents rating of the teaching material in general faceted for each event.}\label{fig:rating-teaching}
\end{figure}

@@

As mentioned above, the following competence areas were identified as relevant for the tutorials: 1. Information and data literacy, 2. Communication and collaboration, 3. Digital content creation and, 5. Problem solving (European Commission, Joint Research Centre et al. 2022). In these competence areas various competences are developed towards an advanced or even specialised level. A more extensive description of the Digital Competence Framework for Citizens in relation to Agent-Based Modelling for archaeologists is available elsewhere (Visser et al. 2024).

@@

\leavevmode\vadjust pre{\hypertarget{ref-visser2024}{}}%
Visser, RM, Lambers, K, Aitchison, K, Jutte, A, Rocks-Macqueen, D, Knaap, L van der, Romanowska, I and Brughmans, T. 2024 \emph{The Digital Competence Framework for Citizens (2.2) and Agent-Based modelling for Archaeologists}. DOI: https://doi.org/\href{https://doi.org/10.5281/zenodo.11046488}{10.5281/zenodo.11046488}.

>>>>

\title{Lessons learned: creating tutorials teaching Agent-Based Modelling for Archaeologists}
\author{}
\date{\vspace{-2.5em}November 13, 2024}

@@

The surveys were analysed using R (R Core Team 2023). The following packages were used for analyses: ggplot2 (Wickham 2016), dplyr (Wickham et al. 2023), tidyr (Wickham, Vaughan and Girlich 2023), forcats(Wickham 2023a), lubridate (Grolemund and Wickham 2011) and stringr (Wickham 2023b). For the sake of reproducibility (Marwick 2017) this paper was written in R Markdown (\url{https://rmarkdown.rstudio.com/}), with data and code available at \url{https://github.com/ABMArchaeologists/ABMA_paper} (Visser et al. 2024a).

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/abm-knowledge-1.pdf}
\caption{\label{fig:abm-knowledge}Respondents knowledge of ABM per event (top), and knowledge of the various softwares for ABM (bottom), before participating in the workshops.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/rating-teaching-1.pdf}
\caption{\label{fig:rating-teaching}Respondents rating of the teaching material (top) and teachers (bottom) in general faceted for each event.}
\end{figure}

@@

As mentioned above, the following competence areas were identified as relevant for the tutorials: 1. Information and data literacy, 2. Communication and collaboration, 3. Digital content creation and, 5. Problem solving (European Commission, Joint Research Centre et al. 2022). In these competence areas various competences are developed towards an advanced or even specialised level. A more extensive description of the Digital Competence Framework for Citizens in relation to Agent-Based Modelling for archaeologists is available elsewhere (Visser et al. 2024b).

@@

\leavevmode\vadjust pre{\hypertarget{ref-visser}{}}%
Visser, RM, Jutte, A, Rocks-Macqueen, D, Romanowska, I, Brughmans, T, Knaap, L van der, Lambers, K, Aitchison, K, Linssen, J and Pažout, A. 2024a \emph{ABMA: Lessons learned. Data and source code}. DOI: https://doi.org/\href{https://doi.org/10.5281/zenodo.14134218}{10.5281/zenodo.14134218}.

\leavevmode\vadjust pre{\hypertarget{ref-visser2024}{}}%
Visser, RM, Lambers, K, Aitchison, K, Jutte, A, Rocks-Macqueen, D, Knaap, L van der, Romanowska, I and Brughmans, T. 2024b \emph{The Digital Competence Framework for Citizens (2.2) and Agent-Based modelling for Archaeologists}. DOI: https://doi.org/\href{https://doi.org/10.5281/zenodo.11046488}{10.5281/zenodo.11046488}.
